### 2024-09-27 paper/paper.qmd
Final commit

---
title: "Trends and Challenges in Toronto’s Homeless Population (2018-2024)"
subtitle: "Middle-Aged Males Leading the Homeless Population with Continued Growth in Chronic Homelessness"
author: 
  - Jianing Li
thanks: "Code and data are available at: https://github.com/JianingLi1225/Toronto_Homeless"
date: today
date-format: long
abstract: "Toronto’s homeless population experienced changes between 2018 and 2024, with key trends in age, gender, and chronic homelessness. Using data from the Toronto Shelter System, this paper examines these trends and the effects of the COVID-19 pandemic. The results show that males aged 35-44 are the largest demographic, and the number of chronically homeless individuals has increased. Understanding these dynamics is essential for developing more effective policies to address homelessness and improve long-term housing stability."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false
#|
#### Workspace setup ####
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(styler)
```

# Introduction

Homelessness is a growing global concern, especially with rapid urbanization and the housing crisis. As Canada's largest city, Toronto faces a growing homeless population [@toronto2024]. In addition to being a socio-economic issue, homelessness intersects with public health and social justice [@healthhomelessness].  However, the existing data often lacks detailed classification, preventing a comprehensive understanding of the needs of specific groups. The present study seeks to fill this gap by analyzing the share and trends of different groups in the homeless population.

Using the "Toronto Shelter System Flows" dataset from the Toronto Open Data Platform [@torontosheltersystem], this paper examines trends in Toronto's homeless population from 2018 to 2024, focusing on age, gender, and chronic homelessness. Our analyses show that homelessness declined during the pandemic and then rebounded. Among all homeless people, males and the 35-44 age group make up the majority, and the number of chronically homeless people continues to grow. These results provide important insights for policymakers.

In this essay, @sec-data-overview explains the background and necessity for using the dataset, followed by an explanation of the dataset's measurements. @sec-data-results shows the changes in the number of homeless individuals, along with the proportions and trends over time for each age group, gender, and chronic homelessness. Graphs and related explanations are provided. @sec-discussion further analyses the information provided in the figures and describes the weaknesses of the study.

# Data {#sec-data}
## Overview {#sec-data-overview}

The dataset used for this analysis is the “Toronto Shelter System Flows” dataset from the Toronto Open Data Platform [@torontosheltersystem]. It focuses on the measurement of homeless individuals' flow within Toronto’s shelter system, which is funded and operated by the City of Toronto. It covers data from 2018 to September 2024, with the homeless count based on shelter admission records, including demographic information such as age and gender. The dataset also provides data on specific groups, including chronically homeless individuals, refugees, youth, single adults, and families. The measurements are collected through the Shelter Management Information System (SMIS) and are updated monthly to ensure that trends in the homeless population are accurately and timely reflected [@torontosheltersystem].

The variables of interest in this analysis vary by age group (Under 16, 16-24, 25-34, 35-44, 45-54, 55-64, 65 and over) and gender (Male, Female, Transgender/Non-Binary), as well as Chronic and All populations. According to @torontosheltersystem, Chronic refers to those who, according to the federal government's definition of chronic homelessness, meet one of the following two criteria: first, stayed in a shelter for at least 180 nights in the past year; and second, stayed in a shelter multiple times in the past three years, totaling at least 546 cumulative nights. All populations refers to the total number of homeless persons covered in the report without any demographic breakdown.

Similar datasets available on the City of Toronto's Open Data Catalogue include Daily Shelter & Overnight Service Occupancy & Capacity and Daily Shelter Occupancy [@opendata]. However, these datasets only provide an overview of shelter usage without categorizing the homeless population using the shelters. As a result, they were not used in this analysis.

All processes are performed through the `R` language [@citeR]. Data is simulated, tested, and cleaned using the `tidyverse` [@tidyverse] packages, which also facilitate downloading data through `opendatatoronto` [@citeopendatatorontoR] . Unneeded variables are removed and renamed using `dplyr` [@dplyr] and `tidyr` [@tidyr] packages, while graphs are plotted with `ggplot2` [@ggplot2]. Code is formatted using `styler` [@styler], and the final PDF report is generated with `knitr` [@knitr] package. The workflow and code are based on guidelines from @tellingstorieswithdata.

## Results {#sec-data-results}

```{r}
#| echo: false
#| warning: false
#| message: false

# Loading the Cleaned Dataset #
cleaned_shelter_data <- read_csv(here::here("data", "analysis_data", "analysis_data.csv"))

# This code was developed with assistance from ChatGPT.
# Convert date strings and extract year
cleaned_shelter_data$Date <- as.Date(paste0("01-", cleaned_shelter_data$Date), format = "%d-%b-%y")
cleaned_shelter_data$year <- as.numeric(format(cleaned_shelter_data$Date, "%Y"))
```

In this section, key trends in the dataset are explored through visualizations, highlighting changes in the total homeless population over time, the distribution across age groups and genders, and the variation in both the number and percentage of the chronic homeless population.

```{r}
#| echo: false
#| warning: false
#| label: fig-total-population
#| fig-cap: "Trend of Homeless Population in Toronto from 2018 to 2024"

# Filter to include only the "All Population" group
# This code was developed with assistance from ChatGPT.
total_population_data <- subset(cleaned_shelter_data, Population_Group == "All Population")

# Load ggplot2 for plotting
# This code is based on Alexander (2023).
ggplot(total_population_data, aes(x = Date, y = Homeless_Count)) +
  geom_line(color = "blue", linewidth = 1.5) +
  labs(
    x = "Date",
    y = "Homeless Population Count"
  ) +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

@fig-total-population illustrates the fluctuation in Toronto's homeless population from 2018 to 2024. The chart shows cyclical rises and falls, with more pronounced declines around 2020 and 2021, followed by a steady increase. The population reaches its highest point in 2024. Throughout this period, the homeless population varies between approximately 7,500 and 11,000 individuals.

```{r}
#| echo: false
#| warning: false
#| label: fig-age-distribution-sidebyside
#| fig-cap: "Homeless Population by Age Group in Toronto"
#| fig-subcap: ["Age Group Distribution (2018-2024)", "Trends in Age Group Proportions Over Time"]
#| layout-ncol: 2

# Summarize data for all years (total homeless population by age group)
# This code was developed with assistance from ChatGPT.
total_data <- cleaned_shelter_data %>%
  summarise(
    Age_Under_16 = sum(Age_Under_16, na.rm = TRUE),
    Age_16_24 = sum(Age_16_24, na.rm = TRUE),
    Age_25_34 = sum(Age_25_34, na.rm = TRUE),
    Age_35_44 = sum(Age_35_44, na.rm = TRUE),
    Age_45_54 = sum(Age_45_54, na.rm = TRUE),
    Age_55_64 = sum(Age_55_64, na.rm = TRUE),
    Age_65_Over = sum(Age_65_Over, na.rm = TRUE)
  )

# Reshape data for plotting
total_data_long <- tidyr::pivot_longer(total_data,
  cols = starts_with("Age"),
  names_to = "Age_Group",
  values_to = "Count"
)

# Modify age group labels for simplicity
total_data_long <- total_data_long %>%
  mutate(Age_Group = case_when(
    Age_Group == "Age_Under_16" ~ "Under 16",
    Age_Group == "Age_16_24" ~ "16-24",
    Age_Group == "Age_25_34" ~ "25-34",
    Age_Group == "Age_35_44" ~ "35-44",
    Age_Group == "Age_45_54" ~ "45-54",
    Age_Group == "Age_55_64" ~ "55-64",
    Age_Group == "Age_65_Over" ~ "65 and over",
    TRUE ~ Age_Group
  ))

# Calculate percentages

total_count <- sum(total_data_long$Count)
total_data_long <- total_data_long %>%
  mutate(Percent = round(Count / total_count * 100))

# Create a pie chart with percentage labels
ggplot(total_data_long, aes(x = "", y = Count, fill = Age_Group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Percent, "%")),
    position = position_stack(vjust = 0.5), size = 4
  ) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Age Group") +
  scale_fill_brewer(palette = "Set2")

# Summarize data by year and age group
yearly_data <- cleaned_shelter_data %>%
  group_by(year) %>%
  summarise(
    Age_Under_16 = sum(Age_Under_16, na.rm = TRUE),
    Age_16_24 = sum(Age_16_24, na.rm = TRUE),
    Age_25_34 = sum(Age_25_34, na.rm = TRUE),
    Age_35_44 = sum(Age_35_44, na.rm = TRUE),
    Age_45_54 = sum(Age_45_54, na.rm = TRUE),
    Age_55_64 = sum(Age_55_64, na.rm = TRUE),
    Age_65_Over = sum(Age_65_Over, na.rm = TRUE)
  )

# Reshape the data for plotting
yearly_data_long <- tidyr::pivot_longer(yearly_data,
  cols = starts_with("Age"),
  names_to = "Age_Group",
  values_to = "Count"
)

# Modify age group labels for simplicity
yearly_data_long <- yearly_data_long %>%
  mutate(Age_Group = case_when(
    Age_Group == "Age_Under_16" ~ "Under 16",
    Age_Group == "Age_16_24" ~ "16-24",
    Age_Group == "Age_25_34" ~ "25-34",
    Age_Group == "Age_35_44" ~ "35-44",
    Age_Group == "Age_45_54" ~ "45-54",
    Age_Group == "Age_55_64" ~ "55-64",
    Age_Group == "Age_65_Over" ~ "65 and over",
    TRUE ~ Age_Group
  ))

# Calculate total homeless population per year and add percentage column
yearly_data_long <- yearly_data_long %>%
  group_by(year) %>%
  mutate(
    Total_Count = sum(Count),
    Percent = Count / Total_Count * 100
  )

# Create a line chart showing the percentage of each age group over time without a title
ggplot(yearly_data_long, aes(x = year, y = Percent, color = Age_Group, group = Age_Group)) +
  geom_line(size = 1.2) +
  labs(x = "Year", y = "Percentage of Homeless Population", color = "Age Group") +
  scale_color_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```

@fig-age-distribution-sidebyside shows the percentage of Toronto's homeless population by age group from 2018 to 2024, along with changes in each group's share over time. Overall, the largest age group is 35-44 years old, accounting for about 21\% of the total population, followed by the 25-34 age group (approximately 18\%), the 45-54 age group (approximately 17\%), and the 55-64 age group (approximately 14\%). The under 16 and 16-24 age groups each account for 12\%. The smallest age group is 65 years and above, making up about 7\% of the total population.

```{r}
#| echo: false
#| warning: false
#| label: fig-gender-distribution-sidebyside
#| fig-cap: "Homeless Population by Gender in Toronto"
#| fig-subcap: ["Gender Distribution (2018-2024)", "Trends in Gender Proportions Over Time"]
#| layout-ncol: 2

# This code was developed with assistance from ChatGPT.
# Summarize data by year for different genders
gender_data <- cleaned_shelter_data %>%
  group_by(year) %>%
  summarise(
    Male = sum(Male_Count, na.rm = TRUE),
    Female = sum(Female_Count, na.rm = TRUE),
    Transgender_Non_Binary = sum(Transgender_Non_Binary_Count, na.rm = TRUE)
  )

# Calculate total population for each gender over the entire period
total_population <- gender_data %>%
  summarise(Male = sum(Male), Female = sum(Female), Transgender_Non_Binary = sum(Transgender_Non_Binary))

# Reshape data for plotting and update Transgender_Non_Binary to Transgender/Non-binary
total_population_long <- pivot_longer(total_population,
  cols = c("Transgender_Non_Binary", "Female", "Male"),
  names_to = "Gender",
  values_to = "Total_Count"
) %>%
  mutate(Gender = recode(Gender, "Transgender_Non_Binary" = "Transgender/Non-binary"))

# Calculate percentages for each gender
total_population_long <- total_population_long %>%
  mutate(Percent = round(Total_Count / sum(Total_Count) * 100, 1))

# Create a pie chart showing the gender distribution
ggplot(total_population_long, aes(x = "", y = Total_Count, fill = Gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Percent, "%")),
    position = position_stack(vjust = 0.5), size = 4, fontface = "bold", color = "gray30"
  ) +
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set2")

# Reshape data for plotting gender proportions over time
gender_data_long <- pivot_longer(gender_data,
  cols = c("Male", "Female", "Transgender_Non_Binary"),
  names_to = "Gender",
  values_to = "Count"
) %>%
  mutate(Gender = recode(Gender, "Transgender_Non_Binary" = "Transgender/Non-binary"))

# Calculate percentages per year
gender_data_long <- gender_data_long %>%
  group_by(year) %>%
  mutate(
    Total_Count = sum(Count),
    Percent = Count / Total_Count * 100
  )

# Create a line chart to show gender distribution over time
ggplot(gender_data_long, aes(x = year, y = Percent, color = Gender, group = Gender)) +
  geom_line(size = 1.2) +
  labs(x = "Year", y = "Percentage of Homeless Population", color = "Gender") +
  scale_color_brewer(palette = "Set2") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
```

@fig-gender-distribution-sidebyside shows the gender distribution of Toronto's homeless population from 2018 to 2024 and how it has changed over time. Males consistently represent the largest proportion, making up approximately 61.8% of the total homeless population, while females account for about 36.8%, and transgender or non-binary individuals makeup around 1.4%. The percentages of males and transgender or non-binary individuals initially increase but begin to decline after 2021, with the opposite trend observed for females. However, these fluctuations are minor, staying within five percentage points.

```{r}
#| echo: false
#| warning: false
#| label: fig-chronic-population
#| fig-cap: "Changes in Chronic and Non-Chronic Homeless Population and Proportion in Toronto (2018-2024)"

# This code was developed with assistance from ChatGPT.
# Summarize data by year for Chronic and All Population
chronic_data <- cleaned_shelter_data %>%
  filter(Population_Group == "Chronic") %>%
  group_by(year) %>%
  summarize(homeless_count = sum(Homeless_Count))

all_population_data <- cleaned_shelter_data %>%
  filter(Population_Group == "All Population") %>%
  group_by(year) %>%
  summarize(homeless_count = sum(Homeless_Count))

# Create Non-Chronic by subtracting Chronic from All Population
non_chronic_data <- all_population_data
non_chronic_data$homeless_count <- non_chronic_data$homeless_count - chronic_data$homeless_count
non_chronic_data$group <- "Non-Chronic"

# Add Chronic to the dataset
chronic_data$group <- "Chronic"
combined_data <- rbind(chronic_data, non_chronic_data) # Combine data

# Ensure 'group' is a factor and order it so that Chronic appears at the top
combined_data$group <- factor(combined_data$group, levels = c("Non-Chronic", "Chronic"))

# Calculate the proportion of Chronic for each year
proportions <- chronic_data %>%
  left_join(all_population_data, by = "year", suffix = c("_chronic", "_all")) %>%
  mutate(percentage = round((homeless_count_chronic / homeless_count_all) * 100, 1))

# Plot the stacked bar chart with Chronic at the top and add percentages
ggplot(combined_data, aes(x = year, y = homeless_count, fill = group)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    data = proportions, aes(x = year, y = homeless_count_chronic, label = paste0(percentage, "%")),
    vjust = 1.5, color = "white", size = 3
  ) + # Add percentage labels in white color
  labs(x = "Year", y = "Homeless Population Count", fill = "Group") +
  theme_minimal() +
  scale_fill_manual(values = c("Chronic" = "blue", "Non-Chronic" = "orange")) +
  scale_x_continuous(breaks = seq(min(combined_data$year), max(combined_data$year), by = 1))
```

@fig-chronic-population displays the changes in the number and proportion of chronic and non-chronic homeless people in Toronto between 2018 and 2024. The number of chronically homeless has been growing over this time, with over 30,000 people in 2018, and by 2023, the number has gone to about 60,000, nearly doubling. The slight drop in numbers in 2024 is due to the fact that the statistics were last updated on September 17, 2024, and it is projected that by the end of the year, the number will have exceeded the 2023 one. Similarly, the percentage of chronically homeless is growing every year, from 29.7% in 2019 to 63.7% in 2024. It should be noted that this percentage grows faster at about 10 percent between 2019 to 2020 and 2023 to 2024.

# Discussion {#sec-discussion}

## Overview of Homeless Trends (2018-2024)
@sec-data-results presents trends in Toronto's homeless population from 2018 to 2024, focusing on key factors such as age, gender, and chronic versus non-chronic conditions. The homeless population fluctuates greatly, with notable impacts in years like 2020 and 2021 (@fig-total-population), with males and the 35-44 age group making up the majority of the homeless population (@fig-age-distribution-sidebyside, @fig-gender-distribution-sidebyside), and chronic homelessness showing an increasing trend (@fig-chronic-population).

## Impact of COVID-19 on Homelessness
The homeless population declined in 2020 and 2021, likely due to COVID-19. To meet social distancing requirements, the Toronto government reduced shelter capacity and opened temporary shelters, including additional spaces in hotels and motels [@may2023scan]. These actions may explain the drop during these years. However, the rebound in homelessness afterward suggests that issues like housing affordability and lack of social support remain unresolved, highlighting the need for sustainable long-term solutions.

## Demographic and Chronic Homelessness Insights

Analyzing age, gender and chronic homelessness provides insight into Toronto's homeless population. Homelessness is concentrated among working-age adults, particularly the 35-44 and 25-34-year-olds, which may be due to challenges they face with employment or housing stability. During the pandemic, there was a sharp fall in homelessness among the under-16s group, which may reflect government support for young people during the pandemic. However, subsequent increases highlight the need for sustained policies.

In terms of gender, males remain in the majority (61.8\%) and only 1.4\% are transgender or non-binary, with little fluctuation in gender breakdown. In the long term, shelters should build facilities that are targeted to the male/female ratio. At the same time, the needs of transgender and non-binary people cannot be ignored, even if they make up only a small percentage of the population.

The increase in the number of chronically homeless people is largely due to the continued rise in rents and house prices, which are far beyond the ability of many low-income residents to pay, leaving them without the means to maintain stable accommodation [@whitzman2023human]. The rise in the proportion of people who are chronically homeless means that current interventions are not sufficient to help people out of homelessness and that the underlying issues need to be further addressed.

## Weaknesses and next steps
A limitation of this analysis is that the data analyzed does not reflect the entire Toronto homeless population. The database used only includes people who use City of Toronto-funded overnight shelter services and fails to reflect people who spend the night outdoors or use other homeless services. According to the most recent Street Needs Assessment, approximately 18% of the homeless population is not included in the data [@torontosheltersystem].

Future directions for improvement should include expanding the data scope of the analysis to include people who spend the night outdoors and use shelters that are not funded by the Government of Toronto, in order to provide a more complete picture of homelessness in Toronto. This will help fill data gaps and support more effective policy decisions.

\newpage

# References